---
title: "Microbial Communities in the Serengeti"
subtitle: ""
author: "AnaBelle Schwarze"
institute: "University of Illinois at Chicago"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
df <- read_csv("../data/correlation_dataframe_L7.csv")
```

```{r, include=F}
style_xaringan(
  title_slide_background_image = "img/green_microbes.jpg"
)
```


## Background:

- Serengeti National Park, Tanzania

- Variables:
    + rainfall
    + soil type (clay, sand, silt)
    + SOM
    + pH
    + minerals (N, P, Fe, Ca)

## Research questions:

- How do microbial communities in the Serengeti react to the environment?

- Which conditions affect them the most?

- Is there a difference between individual phyla?

---

## Cleaning the Data

```{r}
glimpse(df)
```
 
---

## Cleaning the Data

```{r cleaning-data, echo = TRUE, message = FALSE, warning = FALSE}

df <- df %>%
  mutate(
    kingdom = ifelse(str_detect(bacteria, "k__Archaea"), "Archaea", "Bacteria")
  )

df_clean <- df %>%
  mutate(
    kingdom = str_extract(bacteria, "k__[^;]+") %>% str_remove("k__"),
    phylum  = str_extract(bacteria, "p__[^;]+") %>% str_remove("p__")
  )

df_phylum_k <- df_clean %>%
  group_by(kingdom, phylum) %>%
  summarize(across(Silt:N, mean, na.rm = TRUE), .groups = "drop") %>%
  pivot_longer(
    cols = Silt:N,
    names_to = "variable",
    values_to = "correlation"
  )
```

---

## ANOVA

```{r anova-comparison, echo = F}
kingdom_model <- aov(correlation ~ kingdom, data = df_phylum_k)
summary(kingdom_model)

phylum_model <- aov(correlation ~ phylum, data = df_phylum_k)
summary(phylum_model)
```



---

## ANOVA

```{r indiv-anova, echo = FALSE, warning = FALSE}
anova_results <- df_phylum_k %>%
  group_by(variable) %>%
  do(tidy(aov(correlation ~ kingdom, data = .))) %>%
  filter(term == "kingdom") %>%
  select(variable, p.value, statistic) %>%
  arrange(p.value)

anova_results
```

<span style="color: blue;">Significant (< 0.05): P, Ca, clay, Fe, sand, silt, rainfall (slighly)</span>  

<span style="color: red;">Not significant (> 0.05): SOM, pH, N</span> 
---

## Plotting

.center[
```{r plotting-variables, echo = F}

top_vars <- c("P", "Ca", "Clay", "Fe", "Sand", "Silt")

df_phylum_k <- df_phylum_k %>%
  filter(!is.na(phylum))

df_plot <- df_phylum_k %>%
  filter(variable %in% top_vars)

df_plot <- df_plot %>%
  mutate(phylum = fct_reorder(phylum, correlation))

ggplot(df_plot, aes(x = phylum, y = correlation, fill = kingdom)) +
  geom_col(position = "dodge") +
  facet_wrap(~ variable) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Phyla vs. Environmental Variable", 
       y = "Variable", 
       x = "Phylum") +
scale_fill_manual(
    values = c(
      "Bacteria" = "limegreen",
      "Archaea"  = "mediumpurple3"
    ) 
  ) +
theme(
  plot.title = element_text(size = 10, face = "bold"),
  plot.title.position = "plot",
  axis.title = element_text(size = 10),
  axis.text.y = element_text(size = 6),
  strip.text = element_text(size = 10, face = "bold"),
  legend.position = "top",
  legend.title = element_text(size = 9),
  legend.text = element_text(size = 7),
  plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
) +
labs(fill = "Microbial Kingdom")

```
]

---

## Conclusions

- Bacteria and Archaea respond differently to environmental conditions
    - Kingdom-level differences are a driver of diversity in microbial communities
    
- Soil texture and mineral nutrients shape microbial community structure

- Rainfall/moisture plays a secondary role 


